Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-10-13 paper/sections/proof-reconstruction.tex
[ paper ] reviewed to resolve section.

\subsubsection{Splitting a Goal.}
\label{sssec:splitting-goal}

 To prove a goal, \Metis splits the goal into
disjoint cases. This process produces a list of new subgoals, the
conjunction of these subgoals implies the goal. Then, a proof of the
goal becomes in smaller proofs, one refutation for each subgoal.
These subgoals are introduced in the \TSTP derivation with the \strip
inference rule.

@@

In the following, we show how Theorem~\ref{thm:thm-strip} proves the
correctness of the process mentioned above. But first, let us prove some
lemmas for $unshunt$ function presented in Eq.~(\ref{eq:unshunt-fixed}),
and $split$ function from the definition in Eq.~(\ref{eq:split-definition}).

@@

\begin{itemize}
\item $(φ₁ ⇒ (φ₂ ⇒ φ₃))$
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
  \RightLabel{By def.}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
  \RightLabel{∧⇒-to-⇒⇒}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).}
  \end{bprooftree}
\end{equation*}

\item $(φ₁ ⇒ (φ₂ ∧ φ₃))$
\begin{equation*}
\scalebox{0.95}{
($\mathcal{D}_1$)
\begin{bprooftree}
\AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n) ∧ unshunt(φ₁ ⇒ φ₃, n)$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n-1)$}
\RightLabel{lem-unshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$,}
\end{bprooftree}
}
\end{equation*}

\begin{equation*}
(\mathcal{D}_2)
\begin{bprooftree}
\AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂) ∧ unshunt(φ₁ ⇒ φ₃)$}
\RightLabel{∧-proj₂}
\UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n-1)$}
\RightLabel{lem-unshunt,}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$}
\end{bprooftree}
\end{equation*}
Finally, using the theorem, \texttt{⇒∧⇒-to-⇒∧} from~\cite{AgdaProp},
\begin{equation*}
  \texttt{⇒∧⇒-to-⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃).
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}_1$}
  \AxiomC{$\mathcal{D}_2$}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
  \RightLabel{\tt ⇒∧⇒-to-⇒∧.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
  \end{bprooftree}
\end{equation*}

\end{itemize}
\end{itemize}

\end{proof}

Let us now consider the $split$ function presented in
Eq.~(\ref{eq:split-definition}) that yields the conjunction of subgoals that implies the goal in a \Metis \TSTP derivation.

\begin{definition}[split]
\begin{equation}
\label{eq:split-definition}
\mysplit(φ) =
\begin{cases}
unshunt(\mysplit(φ₁)) &\text{if }φ ≡ φ₁ ∧ φ₂\\
∧\ unshunt(φ₁ ⇒ \mysplit(φ₂)), &\\[2.5mm]
unshunt(¬ φ₁ ⇒ \mysplit(φ₂)),
  & \text{if }φ ≡ φ₁ ∨ φ₂\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(φ₂)),
  & \text{if }φ ≡ φ₁ ⇒ φ₂\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(φ₂))
  & \text{if }φ ≡ φ₁ ⇔ φ₂\\
∧\ unshunt(φ₂ ⇒ \mysplit(φ₁)),
  &\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
  & \text{if }φ ≡ ¬ (φ₁ ∧ φ₂)\\[2.5mm]
unshunt(\mysplit(¬ φ₁))
  & \text{if }φ ≡ ¬ (φ₁ ∨ φ₂)\\
∧\ unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂)),
  &\\[2.5mm]
unshunt(\mysplit(φ₁))
  & \text{if }φ ≡ ¬ (φ₁ ⇒ φ₂)\\
∧\ unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
  &\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(¬ φ₂))
  & \text{if }φ ≡ ¬ (φ₁ ⇔ φ₂)\\
∧\ unshunt(¬ φ₂ ⇒ \mysplit(φ₁)),
  &\\[2.5mm]
unshunt(\mysplit(φ₁)),
  & \text{if }φ ≡ ¬ ¬ φ₁\\
⊤,
  & \text{if }φ ≡ ¬ ⊥\\
⊥,
  & \text{if }φ ≡ ¬ ⊤.
\end{cases}
\end{equation}
\end{definition}

The definition above suffers termination problems since it is not
a structural recursion. Therefore, we have reformulated the equation above to solve such a issue following the technique
described in Section~\ref{ssec:structural-recursion}.

Here and below, we refer to $split$ as the function after applying
the process mentioned above.

@@

\begin{proof} We now proceed by induction on the structure of the
formula φ in the definition of Eq.~(\ref{eq:split-definition}).
For a complete proof, we refer the reader to~\cite{AgdaMetis}.\\[3mm]

($φ ≡ φ₁ ∧ φ₂$). We get a proof for each conjunct and using the introduction rule for conjunction, the result follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
  \RightLabel{By def.}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁) ∧ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁)$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ φ₁$.}
  \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
  \RightLabel{By def.}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁) ∧ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
  \RightLabel{\thmsplit}
  \UnaryInfC{$Γ ⊢ φ₂.$}
  \end{bprooftree}
\end{equation*}

\vskip 3mm

($φ ≡ φ₁ ⇒ φ₂$).
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{}
  \RightLabel{assume φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
  \AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
  \RightLabel{By def.}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{weaken φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ , φ₁ ⊢ \mysplit(φ₂,n-1)$}
  \RightLabel{\thmsplit}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
  \RightLabel{⇒-intro}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$.}
  \end{bprooftree}
\end{equation*}
\vskip 3mm

(φ ≡ ¬ (φ₁ ∨ φ₂)). We show $Γ ⊢ ¬ φ₁$ and $Γ ⊢ ¬ φ₂$. Using their conjunction by the introduction rule, and De Morgan Law, the results follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ unshunt(¬ φ₁) ∧ unshunt(¬ φ₁ ⇒ \mysplit(φ₂, n))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ unshunt(¬ φ₁)$}
\RightLabel{lem-unshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁$,}
\end{bprooftree}
\end{equation*}

@@

  \AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
  \RightLabel{By def.}
  \UnaryInfC{$Γ ⊢ unshunt(¬ φ₁) ∧ unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{∧-proj₂}
  \UnaryInfC{$Γ ⊢ unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
  \RightLabel{\thmsplit}
  \UnaryInfC{$Γ ⊢ ¬ φ₂.$}
  \end{bprooftree}
\end{equation*}
\end{proof}

\begin{theorem}[thm-strip]
\label{thm:thm-strip}
 If $s_i$ are the subgoals for the goal φ, and
$\{ Γ ⊢ s_i \}_{i}$, then $Γ ⊢ φ$.
\end{theorem}

@@

  \AxiomC{$\mysplit(φ) ≡ \bigwedge_i s_i $}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ \mysplit(φ) $}

  \RightLabel{\thmsplit}
  \UnaryInfC{$Γ ⊢ φ$.}
\end{bprooftree}
\end{equation*}
Where, the $subst$ theorem is,
\begin{equation}
  \label{eq:substitution-theorem}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ φ$}   \AxiomC{$ψ ≡ φ$}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ ψ$.}
  \end{bprooftree}
\end{equation}
\end{proof}

% % -------------------------------------------------------------------
% \subsubsection{splitting a Conjunct}
% \label{sssec:splitting-a-conjunct}

% The \conjunct rule extracts from a conjunction, one of its conjunct.
% To achieve this, we use projections repeatedly until we finally find
% the conjunct required.

% \begin{align}
%   \begin{split}
%   conjunct(φ, ψ) =
%     \begin{cases}
%       ψ, &\text{if }φ ≡ ψ\\
%       ψ, &\text{if }φ ≡ φ₁ ∧ φ₁\text{ and }conjunct(φ₁, ψ)≡ ψ\\
%       ψ, &\text{if }φ ≡ φ₁ ∧ φ₁\text{ and }conjunct(φ₂, ψ)≡ ψ\\
%       φ, &\text{otherwise.}
%     \end{cases}
%   \end{split}
% \end{align}

% \begin{theorem}[thm-conjunct]
%   \label{thm:thm-conjunct}
%   $Γ ⊢ φ$, and $ψ \in \Prop$ , then $Γ ⊢ conjunct(φ, ψ)$.
% \end{theorem}

% \begin{proof}
% When φ ≡ ψ, from $Γ ⊢ conjunct(φ, ψ)$ by normalization, we get $Γ ⊢ ψ$.
% By using $subst$ theorem Eq.~(\ref{eq:substitution-theorem}), we get the desire conclusion. If the proposition φ is a conjunction, and $ψ ≡ conjunct(φ_{i}, ψ)$
% for some $i = 1,\ 2$, then,

% \begin{equation}
%   \begin{bprooftree}
%   \AxiomC{$Γ ⊢ φ₁ ∧ φ₂$}
%   \RightLabel{∧-proj$_{i}$}
%   \UnaryInfC{$Γ ⊢ φ_{i}$}
%   \UnaryInfC{$Γ ⊢ conjunct(φ_{i}, ψ)$}
%   \AxiomC{$conjunct(φ_{i}, ψ) ≡ ψ$}
%   \RightLabel{subst}
%   \BinaryInfC{$Γ ⊢ ψ$.}
%   \end{bprooftree}
% \end{equation}

% The last statement follows by using the hypothesis and the proof is complete.
% \end{proof}

% % subsubsection splitting-a-conjunct (end)
% % -------------------------------------------------------------------

@@

% \subsubsection{Reordering.}

% Syntactically speaking, showing the logic equivalence between two conjunctive normal forms is a matter of rearrangements of the conjunctions
% but also of the inner disjunctions inside the propositions involved.
% \improvement{Make clear: we are dealing with CNFs, and prove equivalence between two formulas is equivalent to show two trees are  isomorphic.
% And that problem is NP-hard, and for that, we mostly do brute force.}

% We provide some useful theorems to perform the following tasks.

% \begin{itemize}
%   \item Reorder a disjunction like $φ = φ₁ ∨ φ₂ ∨ \cdots$ to match with another disjunction $ψ = ψ₁ ∨ ψ₂ ∨ \cdots$, when $φ$ is a disjunct of $ψ$.
%   \item Reorder a conjunction of disjuncts (CNF), $φ = φ₁ ∧ φ₂ ∧ \cdots$, to match with another CNF formula equivalent $ψ = ψ₁ ∧ ψ₂ ∧ \cdots$.
% \end{itemize}

% First, we define the $rdisj$ function in Eq.~(\ref{eq:rdisj-def}) to convert a disjunction to be a right-associative disjunction, and following
% a similar approach we define the $rconj$ function for conjunctions.

% \begin{align}
%   \begin{split}
%     \label{eq:rdisj-def}
%     rdisj(φ) =
%     \begin{cases}
%     rdisj(φ₁ ∨ (φ₂ ∨ φ₃)),  & \text{if } φ ≡ (φ₁ ∨ φ₂) ∨ φ₃\\
%     φ₁ ∨ rdisj(φ₂),         & \text{if } φ ≡ φ₁ ∨ φ₂\\
%     φ,                      & \text{otherwise.}
%     \end{cases}
%   \end{split}
% \end{align}

% \begin{lemma}[lem-rdisj]
% \label{lem:lem-rdisj}
%   $Γ ⊢ φ$ then $Γ ⊢ rdisj(φ)$.
% \end{lemma}

% The following function, $build_{∨} : \Prop → \Prop → \Prop$ in Eq.~(\ref{eq:build-or}), intends to construct a disjunction from another disjunction, specifically, we want to rearrange the disjuncts in the source formula to match with the target disjunction for syntactic
% equivalence.

% \begin{equation}
%   \begin{split}
%   \label{eq:build-or}
%   &build_{∨}(φ, ψ) =\\
%   &\hspace{3mm}
%   \begin{cases}
%   ψ, &\text{if } φ ≡ ψ\\
%   ψ, &\text{if } ψ ≡ ψ₁ ∨ ψ₂\text{ and } build_{∨}(φ, ψ_{i}) ≡ ψ_{i}\text{ for some }i=1, 2\\
%   φ, &\text{otherwise.}
%   \end{cases}
%   \end{split}
% \end{equation}

% \begin{lemma}[lem-build$_{∨}$]
% \label{lem:lem-build-or}
% $Γ ⊢ φ$, and $ψ\in \Prop$ then $Γ ⊢ build_{∨}(φ)$.
% \end{lemma}

>>>>

\subsubsection{Strip a Goal.}
\label{sssec:strip-a-goal}

To prove a goal, \Metis splits the goal into
disjoint cases. This process produces a list of new subgoals, the
conjunction of these subgoals implies the goal. Then, a proof of the
goal becomes in smaller proofs, one refutation for each subgoal.
These subgoals are introduced in the \TSTP derivation with the \strip
inference rule as we show with the following excerpt.

@@

To prove the correctness of the process mentioned above, we state
the Theorem~\ref{thm:thm-strip}. But first, let us prove some
lemmas for $unshunt$ function presented in Eq.~(\ref{eq:unshunt-fixed}),
and $split$ function from the definition in Eq.~(\ref{eq:split-definition}).

@@

\begin{itemize}
\item $(φ₁ ⇒ (φ₂ ⇒ φ₃))$
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
  \RightLabel{∧⇒-to-⇒⇒.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃)}
  \end{bprooftree}
\end{equation*}

\item $(φ₁ ⇒ (φ₂ ∧ φ₃))$
\begin{equation*}
  \scalebox{0.95}{
  ($\mathcal{D}_1$)
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n) ∧ unshunt(φ₁ ⇒ φ₃, n)$}
    \RightLabel{∧-proj₁}
    \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n)$}
    \RightLabel{axiom}
    \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n-1)$}
    \RightLabel{lem-unshunt,}
    \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
  }
\end{equation*}

\begin{equation*}
  (\mathcal{D}_2)
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂) ∧ unshunt(φ₁ ⇒ φ₃)$}
    \RightLabel{∧-proj₂}
    \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n)$}
    \RightLabel{axiom}
    \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n-1)$}
    \RightLabel{lem-unshunt.}
    \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$}
    \end{bprooftree}
\end{equation*}

Finally, using the theorem \texttt{⇒∧⇒-to-⇒∧} from~\cite{AgdaProp},
\begin{equation*}
  \texttt{⇒∧⇒-to-⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃),
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}_1$}
  \AxiomC{$\mathcal{D}_2$}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
  \RightLabel{\tt ⇒∧⇒-to-⇒∧.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
  \end{bprooftree}
\end{equation*}

\end{itemize}
\end{itemize}
\end{proof}

Let us now consider the $split$ function presented in
Eq.~(\ref{eq:split-definition}) that yields the conjunction of subgoals that implies the goal of the problem in the \Metis \TSTP derivations.

\begin{definition}[split]
\label{def:split}

\begin{equation}
\label{eq:split-definition}
\mysplit(φ) =
\begin{cases}
unshunt(\mysplit(φ₁)) &\text{if }φ ≡ φ₁ ∧ φ₂\\
∧\ unshunt(φ₁ ⇒ \mysplit(φ₂)), &\\[2.5mm]
unshunt(¬ φ₁ ⇒ \mysplit(φ₂)),
  & \text{if }φ ≡ φ₁ ∨ φ₂\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(φ₂)),
  & \text{if }φ ≡ φ₁ ⇒ φ₂\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(φ₂))
  & \text{if }φ ≡ φ₁ ⇔ φ₂\\
∧\ unshunt(φ₂ ⇒ \mysplit(φ₁)),
  &\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
  & \text{if }φ ≡ ¬ (φ₁ ∧ φ₂)\\[2.5mm]
unshunt(\mysplit(¬ φ₁))
  & \text{if }φ ≡ ¬ (φ₁ ∨ φ₂)\\
∧\ unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂)),
  &\\[2.5mm]
unshunt(\mysplit(φ₁))
  & \text{if }φ ≡ ¬ (φ₁ ⇒ φ₂)\\
∧\ unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
  &\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(¬ φ₂))
  & \text{if }φ ≡ ¬ (φ₁ ⇔ φ₂)\\
∧\ unshunt(¬ φ₂ ⇒ \mysplit(φ₁)),
  &\\[2.5mm]
unshunt(\mysplit(φ₁)),
  & \text{if }φ ≡ ¬ ¬ φ₁\\
⊤,
  & \text{if }φ ≡ ¬ ⊥\\
⊥,
  & \text{if }φ ≡ ¬ ⊤.
\end{cases}
\end{equation}
\end{definition}

The definition above suffers termination problems since it is not
a structural recursion. Therefore, we have reformulated the equation above to solve such a issue following the technique
described in Section~\ref{ssec:structural-recursion}.
Here and below, we refer to $split$ as the function after applying
the process mentioned above.

@@

\begin{proof} We now proceed by induction on the structure of the
formula $φ$ in the definition of Eq.~(\ref{eq:split-definition}).
For a complete proof, we refer the reader to~\cite{AgdaMetis}.\\[3mm]

\begin{itemize}
\item $(φ ≡ φ₁ ∧ φ₂)$. We get a proof for each conjunct and using the introduction rule for conjunction connective, the result follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁) ∧ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁)$}
  \RightLabel{lem-unshunt.}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁) ∧ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
  \RightLabel{\thmsplit.}
  \UnaryInfC{$Γ ⊢ φ₂$}
  \end{bprooftree}
\end{equation*}

\item $(φ ≡ φ₁ ⇒ φ₂)$
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{}
  \RightLabel{assume φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
  \AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{weaken φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ , φ₁ ⊢ \mysplit(φ₂,n-1)$}
  \RightLabel{\thmsplit}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
  \RightLabel{⇒-intro.}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
\end{equation*}

\item $(φ ≡ ¬ (φ₁ ∨ φ₂))$. We show that $Γ ⊢ ¬ φ₁$ and $Γ ⊢ ¬ φ₂$.
The result follows by introducing the conjunction and applying after
De Morgan Law.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{axiom}
\UnaryInfC{$Γ ⊢ unshunt(¬ φ₁) ∧ unshunt(¬ φ₁ ⇒ \mysplit(φ₂, n))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ unshunt(¬ φ₁)$}
\RightLabel{lem-unshunt,}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}
\end{bprooftree}
\end{equation*}
\end{itemize}
\end{proof}

@@

  \AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(¬ φ₁) ∧ unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{∧-proj₂}
  \UnaryInfC{$Γ ⊢ unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
  \RightLabel{\thmsplit.}
  \UnaryInfC{$Γ ⊢ ¬ φ₂$}
  \end{bprooftree}
\end{equation*}

\begin{lemma}[subst]
  \label{lem:subst}
  Substitution theorem.
\begin{equation}
  \label{eq:substitution-theorem}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ φ$}   \AxiomC{$ψ ≡ φ$}
  \RightLabel{subst.}
  \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation}
\end{lemma}

\begin{theorem}[thm-strip]
\label{thm:thm-strip}
 If $s_i$ are the subgoals of the goal $φ$, and
 $\{ Γ ⊢ s_i \}_{i}$, then $Γ ⊢ φ$.
\end{theorem}

@@

  \AxiomC{$\mysplit(φ) ≡ \bigwedge_i s_i $}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ \mysplit(φ)$}
  \RightLabel{\thmsplit.}
  \UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
\end{equation*}
\end{proof}

% -------------------------------------------------------------------
\subsubsection{Splitting a Conjunct.}
\label{sssec:splitting-a-conjunct}

The \conjunct rule extracts from a
conjunction one of its conjuncts; it is a generalization of the
projection rules for the conjunction connective as the
following \TSTP excerpt shows.

\begin{verbatim}
fof(normalize_1, plain, p & q & (r | ~ p),
  inference(canonicalize, [], [x])).
fof(normalize_2, plain, q,
  inference(conjunct, [], [normalize_1])).
fof(normalize_2, plain, (r | ~ p),
  inference(conjunct, [], [normalize_1])).
\end{verbatim}

\begin{definition}[conjunct]
\begin{align}
  \begin{split}
  &conjunct : \Prop \to \Prop \to \Prop\\
  &conjunct(φ, ψ) =\\
  &\hspace{3mm}
  \begin{cases}
      ψ, &\text{if }φ ≡ ψ\\
      ψ, &\text{if }φ ≡ φ₁ ∧ φ₁\text{ and }ψ ≡ conjunct(φ₁, ψ)\\
      ψ, &\text{if }φ ≡ φ₁ ∧ φ₁\text{ and }ψ ≡ conjunct(φ₂, ψ)\\
      φ, &\text{otherwise.}
    \end{cases}
  \end{split}
\end{align}
\end{definition}

\begin{theorem}[thm-conjunct]
  \label{thm:thm-conjunct}
  $Γ ⊢ φ$, and $ψ  : \Prop$ , then $Γ ⊢ conjunct(φ, ψ)$.
\end{theorem}
~\begin{proof}\hspace{2cm}\\
\begin{itemize}
  \item For the case, $φ ≡ ψ$, $Γ ⊢ conjunct(φ, ψ)$ normalizes to $Γ ⊢ ψ$.
Then, we get the desire conclusion by applying the $subst$ theorem.
  \item If the proposition $φ$ is a conjunction, and we can get $ψ ≡ conjunct(φ_{i}, ψ)$ for some $i = 1,\ 2$, then,

\begin{equation}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ φ₁ ∧ φ₂$}
  \RightLabel{∧-proj$_{i}$}
  \UnaryInfC{$Γ ⊢ φ_{i}$}
  \UnaryInfC{$Γ ⊢ conjunct(φ_{i}, ψ)$}
  \AxiomC{$ψ ≡ conjunct(φ_{i}, ψ)$}
  \RightLabel{subst.}
  \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation}
\item The last case follows trivially by using the same hypothesis.
\end{itemize}
\end{proof}

% subsubsection splitting-a-conjunct (end)
% -------------------------------------------------------------------

@@

\subsubsection{Reordering.}
Logic equivalence between propositions is a major issue to justify
prover reasoning steps. Since we left out semantics to treat only the
syntax aspects of the propositions, our approach shows that equivalence
by converting propositions to their conjunctive normal forms and showing equivalence by reordering the conjunctions and the inner disjunctions.
In this section, we provide some theorems to perform such reordering tasks, the goals are the following.

\begin{itemize}
  \item Reorder the disjunction $φ = φ₁ ∨ φ₂ ∨ \cdots φₙ$ to match with $ψ = ψ₁ ∨ ψ₂ ∨ \cdots ∨ ψₙ$, when for all $i = 1, \cdots, n$, exists at least one $ψⱼ$ such that $φᵢ ≡ ψⱼ$ for $j = 1, \cdots, n$.
  \item Reorder a CNF formula, $φ = φ₁ ∧ φ₂ ∧ \cdots$, to match with another CNF formula $ψ = ψ₁ ∧ ψ₂ ∧ \cdots$ when for all $i = 1, \cdots, n$, exists at least one $ψⱼ$ such that $φᵢ ≡ ψⱼ$ for $j = 1, \cdots, n$.
\end{itemize}

First, we define the $rdisj$ function in Eq.~(\ref{eq:rdisj-def}) to convert a disjunction in its right-associative form.
\begin{definition}[rdisj]
\begin{align}
  \begin{split}
    \label{eq:rdisj-def}
    &rdisj(φ) : \Prop \to \Prop\\
    &rdisj(φ) =
    \begin{cases}
    rdisj(φ₁ ∨ (φ₂ ∨ φ₃)),  & \text{if } φ ≡ (φ₁ ∨ φ₂) ∨ φ₃\\
    φ₁ ∨ rdisj(φ₂),         & \text{if } φ ≡ φ₁ ∨ φ₂\\
    φ,                      & \text{otherwise.}
    \end{cases}
  \end{split}
\end{align}
\end{definition}

\begin{lemma}[lem-rdisj]
\label{lem:lem-rdisj}
  $Γ ⊢ φ$ then $Γ ⊢ rdisj(φ)$.
\end{lemma}
Following a similar way to define $rdisj$, we define the $rconj$ function,
to convert a conjunction into its right-associative form.

The $build_{∨}$ function in Eq.~(\ref{eq:build-or}) function intends to construct a disjunction from another disjunction, specifically, it rearranges the disjuncts in the source formula to match with the target disjunction formula as we described above.

\begin{definition}[build$_{∨}$]
\begin{equation}
  \begin{split}
  \label{eq:build-or}
  &build_{∨} : \Prop → \Prop → \Prop\\
  &build_{∨}(φ, ψ) =\\
  &\hspace{3mm}
  \begin{cases}
  ψ, &\text{if } φ ≡ ψ\\
  ψ, &\text{if } ψ ≡ ψ₁ ∨ ψ₂\text{ and } build_{∨}(φ, ψ_{i}) ≡ ψ_{i}\text{ for some }i=1, 2\\
  φ, &\text{otherwise.}
  \end{cases}
  \end{split}
\end{equation}
\end{definition}
